---
title: "output_plots"
format: html

execute: 
  output: true
  echo: false
---

```{r}
library("tidyverse")
library("ggplot2")
library("sf")
library("prismatic")
library("zoo")
library("dataRetrieval")

#Bringing in data
sms <-readNWISdv(siteNumbers = "08170000", 
                 parameterCd = "00060") 
cs <-readNWISdv(siteNumbers = "08168710", 
                parameterCd = "00060") 
df<-bind_rows(sms,cs)|> renameNWISColumns(df)

gages<- df|> 
  mutate( gage_name = case_when(
    site_no=="08170000"~"San Marcos Springs", 
    site_no=="08168710"~"Comal Springs"), 
    year = year(Date), 
    month = month(Date), 
    day = day(Date))|> 
  rename(cfs = Flow)

gages <- gages|>
  arrange(Date)|>
  mutate(
    avg_30_prior = rollmean(cfs,30,na.pad = TRUE, align = "right"),
    avg_60_prior = rollmean(cfs,60,na.pad = TRUE, align = "right"),
    avg_10_prior = rollmean(cfs,10,na.pad = TRUE, align = "right"))|>
  select(site_no,gage_name,Date,year,month,day,cfs,avg_10_prior,avg_30_prior,avg_60_prior)

(sucralose_raw <- readr::read_csv("C:/Users/mgingerich/OneDrive - Edwards Aquifer Authority/Desktop/Sucralose_Project/data/sucralose_simplified.csv", col_types = cols(
  Date = col_date("%m.%d.%Y"),)))

sucralose <- sucralose_raw|>
  mutate(add_column(sucralose_raw,
    gage_name = case_when(
      str_detect(sys_sample_code, "CM") ~ "Comal Springs",
      str_detect(sys_sample_code, "SM") ~ "San Marcos Springs")))|>
  relocate("gage_name",.before="Date")

sucralose_no_dup <- subset(sucralose,duplicate == "N")
sucralose_no_dup_blank <- subset(sucralose, blank == "N")
false_pos <- sucralose|>
  filter(detect_flag == "Y"& blank == "Y")

stream_sucralose <- sucralose|>
  left_join(gages, join_by(Date,gage_name))|>
  select(Date,gage_name,cfs,avg_10_prior,avg_30_prior,avg_60_prior,result_value,lab_qualifiers,detect_flag)
```

### Springflow plots

Showing average springflow (cfs) over 4 time periods by gage

```{r}
theme(theme_bw())
avg <- gages|>
  group_by(gage_name)|>
  add_column(
    label = "avg")|>
    ungroup()

avg_25 <- gages|>
  group_by(gage_name)|>
  filter(year =="2025")|>
  add_column(
    label = "avg_25")|>
  ungroup()

avg_3yr <- gages|>
  group_by(gage_name,year)|>
  filter(year %in% c("2025","2024","2023"))|>
  add_column(
    label = "avg_3yr")|>
  ungroup()


avgs <- bind_rows(avg_25,avg_3yr,avg)|>
  select(gage_name,cfs,label)

avgs$label <- factor(avgs$label,
                     levels = c("avg_25","avg_3yr","avg")) 

ggplot(avgs, aes(gage_name,cfs, color = label))+
  geom_boxplot(aes(gage_name,cfs, color = label, fill = after_scale(clr_desaturate(clr_lighten(color,.8),.8))),linewidth = 0.65)+
  stat_summary(fun=mean, geom = "point",shape = 4, position = position_dodge2(
    width = 0.75, preserve = "single"))+ #marks the mean on the boxplot with an x
  labs(
    title = "Average Springflow over Time",
    subtitle = "Comal Springs and San Marcos Springs",
    y = "Springflow (cfs)",
    x = "")+
  scale_color_discrete(name = element_blank(),
                       labels = c("2025","3-year","Long-term"))+
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_text(margin = margin(r = 10)))
```

Showing average flow for each year with cfs data for comal and sm

```{r}
theme(theme_bw())
mean_flow_gage_year <- gages |>
  group_by(gage_name,year)|>
  summarise(mean_flow = mean(cfs,na.rm=TRUE))|>
  ungroup()
mean_flow_gage_year

mean_flow_plot <- ggplot(mean_flow_gage_year)+
  geom_line(aes(year,mean_flow, color = gage_name))+
  geom_smooth(aes(year,mean_flow),
              color = "darkgrey",linetype = "dashed", se = FALSE)+
  labs(
    title = "Annual Mean Flow by Gage",
    x = "Year",
    y = "Mean Flow (cfs)"
  )+
  theme(legend.title = element_blank(),
        legend.position = "none")+
  facet_wrap(~gage_name)
mean_flow_plot
```
```{r}
theme_set(theme_bw())

springflow_plot <- gages|>
  ggplot()+
  geom_line(aes(x = Date, y = cfs, color = gage_name), linewidth = 1)+
  scale_color_brewer(palette = "Paired")+
  labs(
    title = "",
    x = "Date",
    y = "Springflow (cfs)")+
  theme(
    legend.title = element_blank(),
    legend.position = "bottom")

springflow_plot +inset_element(
  cfs_plot,
  left = 0.65,
  bottom = 0.55,
  right = unit(1, 'npc')-unit(1.55, 'cm'),
  top = unit(1, 'npc')-unit(0.02, 'cm')
)
ggsave("springflow_plot_blue_insetboxplot.png")
```
Showing daily flow by gage, with dotted lines representing flow goal min and max for each gage

```{r}
library("patchwork")
theme(theme_bw())

cfs_plot_cs <- gages|>
  filter(gage_name =="Comal Springs")|>
  ggplot()+
  geom_line(aes(x = Date, y = cfs), color = "salmon", lwd = 0.75)+
  facet_wrap(~gage_name)+
  theme(legend.position = "none",
        axis.title.x = element_blank())

cfs_plot_sms <- gages|>
  filter(gage_name =="San Marcos Springs")|>
  ggplot()+
  geom_line(aes(x = Date, y = cfs), color = "turquoise", lwd = 0.75)+
  facet_wrap(~gage_name)+
  theme(legend.position = "none",
        axis.title.x = element_blank())

cfs_plot_cs + geom_hline(yintercept = c(225,30), linetype = "dashed", color = c("springgreen3","red"))+
cfs_plot_sms + geom_hline(yintercept = c(140,80), linetype = "dashed", color = c("springgreen3", 'red'))+ plot_annotation('Daily Flow by Gage', caption = 'Source: USGS',theme=theme(plot.title=element_text(hjust=0.5))) #adding dashed lines that represent the flow goals for the respective spring
```

```{r}
#summary table for springflow
library("gt")

spring_stats|>
  rename(
    "Gage" = gage_name,
    "Mean (cfs)" = mean_cfs,
    "Standard Deviation (cfs)"= stdv_cfs,
    "Min (cfs)" = min_cfs,
    "Max (cfs)" = max_cfs,
    "Median (cfs)" = median_cfs)|>
  gt()
#gtsave("spring_stats.html")

```

How can I determine the avg and range of suc levels detected in the edwards?

```{r}
theme(theme_bw())
cfs_plot <- gages|>
  ggplot()+
  geom_boxplot(aes(gage_name,cfs, color = gage_name, 
                   fill = after_scale(clr_desaturate(
                     clr_lighten(color,.8),.8))),linewidth = 0.5, outlier.alpha = 0)+
  scale_color_brewer(palette = "Paired")+
  labs(
    title = "",
    x = "",
    y = "")+
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.margin = unit(c(0.01,0.01,0,0.01), "cm")) #top, right, bottom, left
cfs_plot
```
### Sucralose plots
Do suc levels vary with streamflow from previous days?

```{r}
library("patchwork")
theme(theme_bw())
plot1 <- stream_sucralose|>
  filter(detect_flag == "Y")|> #removing non-detects
ggplot()+
  geom_point(aes(avg_30_prior,result_value, color = gage_name,shape=gage_name), na.rm = TRUE)+
  scale_color_brewer(palette = "Paired")+
  theme(legend.position = "none",)+
  labs(
    x = "30-day Prior Avg Flow (cfs)",
    y = "Sucralose (ng/L)")

plot2 <- stream_sucralose|>
  filter(detect_flag == "Y")|>
  ggplot()+
  geom_point(aes(avg_10_prior, result_value, color = gage_name,shape = gage_name),na.rm = TRUE)+
  scale_color_brewer(palette = "Paired")+
  theme(legend.position = "none")+
  labs(
  title = "Sucralose Levels by Flow",
  x = "10-day Prior Avg Flow (cfs)",
  y = "Sucralose (ng/L)")

plot3 <- stream_sucralose|>
  filter(detect_flag == "Y")|>
  ggplot()+
  geom_point(aes(avg_60_prior, result_value, color = gage_name,shape=gage_name), na.rm = TRUE)+
  scale_color_brewer(palette = "Paired")+
  theme(legend.position = "bottom",
        legend.margin = margin(l=220),
        legend.title = element_blank())+
  labs(
    x = "60-day Prior Avg Flow (cfs)",
    y = "Sucralose (ng/L)")

plot4 <- stream_sucralose|>
  filter(detect_flag == "Y")|>
  ggplot()+
  geom_point(aes(cfs,result_value,color=gage_name,shape=gage_name),na.rm=TRUE)+
  scale_color_brewer(palette = "Paired")+
  theme(legend.position = "none")+
  labs(
    x = "Day-of Flow (cfs)",
    y = "Sucralose (ng/L)")


plot2+plot1+plot3+plot4+
  plot_layout(design = layout)
ggsave("suc_flow_blue.png")
```

How have suc averages changed over time at each gage?

```{r}
theme(theme_bw())
p1 <- stream_sucralose|>
ggplot()+
  geom_point(aes(Date,result_value, color = gage_name), na.rm = TRUE)+
  geom_smooth(aes(Date,result_value, color = gage_name),se=FALSE)+
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        legend.margin = margin(l=220))+
  labs(
    title = "Sucralose Levels by Gage and Date",
    subtitle = "Comal and San Marcos Springs",
    x = "",
    y = "Sucralose (ng/L)")

p2 <- stream_sucralose|>
  ggplot()+
  geom_boxplot(aes(gage_name,result_value, color = gage_name, 
                   fill = after_scale(clr_desaturate(
                     clr_lighten(color,.8),.8))),linewidth = 0.65)+
  labs(x = "",
    y = "Sucralose (ng/L)")+
  theme(legend.position = "none")

p1+p2
```


```{r}
#summary statistics tables for results section
#install.packages("gt")
#install.packages("janitor")
library("janitor")
library("gt")

suc_stats_wide <- suc_summary_stats|>
  rename(
    "Gage" = gage_name,
    "n Total" = n_tot,
    "n SUC Detected" = num_detect,
    "Mean (ng/L)" = mean_suc,
    "Min (ng/L)" = min_suc,
    "Max (ng/L)" = max_suc,
    "Median (ng/L)" = median_suc,
    "Mean Detection Limit (ng/L)" = mean_detect)|>
  select("Gage","n Total", "n SUC Detected","Mean (ng/L)","Min (ng/L)","Max (ng/L)", "Median (ng/L)","Mean Detection Limit (ng/L)")

tab2 <- gt(suc_stats_wide)|>
  fmt_number(decimals = 2)
tab2
#gtsave(tab2, filename = "suc_summary_table.png")
```


### Suc dataset analysis

Do duplicate measurements match their original?

```{r}
theme(theme_bw())
duplicates <- sucralose|>
  filter(gage_name %in% c("San Marcos Springs","Comal Springs"),
         duplicate == "Y" | lead(duplicate) == "Y")|>
  mutate(label = case_when(
    str_detect(duplicate, "N")~"Original",
    str_detect(duplicate, "Y")~"Duplicate")) #creating new column for duplicate identification

dup_mean <- duplicates|>
  group_by(Date)|>
  mutate(
    avg = mean(result_value))|>
  slice(1:1)|> 
  ungroup() #calculates the avg between the original and duplicate measurements and slice the 1st value to plot a mean for each pair

dup_plot <- duplicates|>
  ggplot()+
  geom_point(data = duplicates, aes(x = Date, y = result_value, color = label, shape = label),size = 3)+
  geom_point(data = dup_mean, aes(x = Date, y = avg), shape = 4)+
  labs(
    title = "Analysis of Duplicate Sucralose Measurements",
    subtitle = "Original and Duplicate Results",
    x = "Date",
    y = "Sucralose (ng/L)",
    caption = "*cross represents the average for each original/duplicate pair")+
  theme(legend.title = element_blank())
dup_plot
```

Do DI blanks return the reporting detection limit?

```{r}
theme(theme_bw())
blanks_plot <- false_pos|>
  ggplot()+
  geom_point(aes(sys_sample_code,result_value, 
                 color = gage_name, shape = gage_name), size = 3)+
  geom_point(aes(sys_sample_code,reporting_detection_limit), shape = 4)+
  labs(
    title = "Analysis of DI Blank Measurements",
    x = "Date",
    y = "Sucralose (ng/L)",
    caption = "*cross represents the reporting detection limit")+
  theme(legend.title = element_blank())+
  theme(legend.position = "bottom")
blanks_plot
```
```{r}
data <- data.frame(
  x = c("Comal Springs","San Marcos Springs","WW influence", "Recreational influence","WW Effluent"),
  y = c(9.65,16.0,680,76.5,2400))

data$x <- factor(data$x, levels = c("Comal Springs","San Marcos Springs","Recreational influence","WW influence","WW Effluent"))
                 
theme_set(theme_bw())

comp_plot <- ggplot()+
  geom_col(aes(x=data$x, y=y, fill = data$x))+
  geom_label(position = position_nudge(y = 100), aes(x=data$x,y=y,label = y), size = 3.25, label.size = 0)+
  labs(x = "",
       y = "Median SUC (ng/L)",
       title = "Sucralose Concentrations with Varying Contamination")+
  scale_fill_manual(
    values = c("Comal Springs" = "darkslateblue", 
               "San Marcos Springs" = "cornflowerblue",
               "Recreational influence" = "darkslategray2",
               "WW influence" = "darkturquoise",
               "WW Effluent" = "darkcyan"))+
  theme(axis.text.x = element_blank(),
        legend.title = element_blank(),
        axis.title.y = element_text(vjust = 1),
        legend.position = "bottom")

comp_plot
ggsave(filename = "SUC_comparison_plot_titled.png")
```